Distributional Trees and Forests

### 2017-11-27 pkg/disttree/inst/draft/draft.Rnw
small corrections
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/partykit@1758 255f4249-2263-4417-b1ea-8498779e27a7

Former-commit-id: 81caa5ae1744d544d82350e6d40ef169a35eb40a

\SweaveOpts{eval = TRUE}
\section{Introduction}
A typical application in probabilistic forecasting in meteorology is the analysis of precipitation data. The main interest lies in estimating the probability of precipitation as well as predicting the expected amount of precipitation. In this research area various methods have been developed and tested in the last decades. (Multiple) Linear models such as ensemble model output statistics (EMOS) were introduced more than a decade ago (\citealt{gneiting2005calibrated}) and are still common and widely used. Despite their simplicity and interpretability, the limitation to modeling only linear effects has risen the interest in applying more flexible models. This feature is provided by generalized additive models (GAM) which can also model non-linear effects. However, often only the expectation of the underlying distribution is estimated. To obtain probabilistic predictions, generalized additive models for location, scale, and shape (GAMLSS, \citealt{stasinopoulos2005}) offer a framework in which each distribution parameter can be modeled separately by its own GAM. In Figure~\ref{Devel_parmod} the development of parametric models starting from simple linear models (LM) and ranging to GAMLSS is illustrated.\\
These flexible parametric methods provide well fitting models when applied on precipitation data in a rather smooth and homogeneous setting, e.g. areas where no abrupt changes in meteorological/climatological events appear. However, for example in complex terrain in mountainous regions various factors can change rapidly and their effects are difficult to be represented by a smooth parametric model. In this situation a different approach which can deal with these abrupt changes would be to apply tree structured models such as regression trees (\citealt{loh2011classification}) or random forests (\citealt{breiman2001random}). Moreover, these models are particularly advantageous in case of many covariates with unknown effects or interactions as the variable selection is done automatically within the model fitting process. Therefore, no additional effort or knowledge for the specification of the model is required in advance. However, even though their main advantage of being able to model non-additive effects and interactions without any necessary prespecification can play to their strength, regression trees are not an adequate choice if (additional) smooth effects appear as well. In this situation a random forest provides a good compromise as it combines a set of trees such that the steps imposed by the trees are smoothed out in case of smooth effects while abrupt changes are still captured by the tree structure of the model.\\
For these reasons, regression trees and particularly random forests provide valuable features which can contribute to improving models for probabilistic forecasting. But, up to now, no general distributional framework is available for these methods. Therefore, we are going to introduce distributional trees and forests which embed the idea of flexible parametric modeling into the framework of regression trees and random forests. The developments in the field of tree-structured models together with the novel approaches of distributional trees and forests are illustrated in Figure~\ref{Devel_treeforest}.\\

@@

After explaining the novel methodology in Section~\ref{Methodology} of this paper the focus will then be put on an application in the field of probabilistic forecasting in Section~\ref{ProbabilisticForecasting}. The new method is applied on precipitation data in a mountainous region with a large number of numerical weather prediction quantities and the results are compared to those of other methods commonly used in this research area. In a first setting only one observation station is taken into consideration. The models are trained on data from 24 years and probabilistic predictions of the consecutive 4 years are validated. A brief outlook on the results of this setting is given by Figure~\ref{axams_onestation}. For the 24th of July of 2009, 2010, 2011 and 2012 the density functions of the distributions estimated by a forest model are plotted together with the corresponding observations. This illustrates that well fitting predicted expected values are provided by this model as they are close to the observed values.
Beside the expected values the predictions provided by distributional models offer a wide range of information as they specify the whole distribution and therefore enable further methods of inference and analysis. \\

@@

where $\phi$ and $\Phi$ are the probability density function and the distribution function of the standard normal distribution $\mathcal{N}(0,1)$. Next to the censored normal distribution the censored logistic distribution is also often chosen for the modeling of precipitation amounts. In \citealt{Gebetsberger2017} this choice has been made in order to be able to deal with heavy tails which often appear in precipitation data. Another frequent application of probabilistic forecasting in meteorology is the modeling of temperature. In this case a normal distribution is often selected.\\

@@

In that way, a whole distribution is specified with all its features, including location, scale and shape. However, fitting one global model to the whole data set might be too much of a generalization and therefore not represent the data and its features in a reasonable way. Moreover, if covariates are available it is desirable to include them into the model as they might provide important additional information. In our application on precipitation data a wide range of covariates are available from numerical ensemble predictions considering precipitation, temperature, sea level pressure and other meteorological quantities which might be of interest and have an influence on the predictions. \\

@@

However, the restriction to additive models remains. Therefore, a different approach of how to include covariates that also allows for non-additive effects and interactions is considered in the next subsection of this article. In particular, the covariates can be used to separate the data into more homogeneous subgroups. Doing so after fitting a global model to the whole data set (without considering any covariates) and then fitting a local model to each subgroup should improve the model remarkably. This procedure of splitting a data set into subgroups can be done by applying a tree algorithm.\\

@@

While the clear structure of such a tree model is advantageous in interpreting and analyzing it, the abrupt changes are often too rough and impose steps on the model even if the true underlying effect is smooth. In these situations bagging can solve the problem by combining an ensemble of trees such that wrongly detected abrupt changes of the model are turned into almost smooth transitions. In that way, all kinds of effects can be modeled, which is is also one the of the main advantages of forest models (taking bagging one step further as the variance is reduced by sampling the covariates considered for the splitting process in each node). \\
\\
Common forest models such as random forests consist of trees only considering the mean of the distribution of the response variable. For that reason, changes in other parameters can only be detected if they are connected to changes in the location parameter, e.g. if they depend on the location parameter. Therefore, building a forest of distributional trees which can also capture separate changes in each of the distribution parameters leads to an improvement of the model. This valuable feature of detecting separate changes is also provided by GAMLSS as each parameter is modeled separately. However, in this case this separation can also be a disadvantage as it hinders the capturing of interactions unless they are specified explicitly by the user. These two cases provide once again a good reason to combine the two frameworks of flexible parametric modeling and tree structured models in order to profit from their advantages and compensate possible disadvantages. This is the pursued goal in introducing distributional forests.

\subsection{Distributional Forest}
\label{distforest}
A distributional forest consists of an ensemble of distributional trees $\{t_s\}_{s=1,\ldots,T}$. Each of these trees is built on a different data set which can either be a bootstrap or a subset of the original data set. Moreover, in each node a subset of the partitioning variables is chosen randomly as candidates for splitting variables. In that way the correlation among the resulting trees is reduced and therefore also the variance of the model (\citealt{breiman2001random}).\\
\\
Forest models differ mainly in the way how the trees are combined to make predictions. In this case the forest itself is only used to evaluate neighborhood relationships among the observations which can then be included in the fitting process.\\
Contrary to distributional trees, in distributional forests the whole learning data set can be used to fit the distributional model and specify the predicted distribution for a new set of covariates. This difference is represented in the weight vector as the tree weights $\bold{\omega}^{tree}$ containing only zeros and ones are now replaced by integer valued forest weights \\
$\bold{\omega}^{forest} = (\omega^{forest}_1, \ldots, \omega^{forest}_n)$.\\
For a new set of covariates $\bold{z}^{\ast} = (z^{\ast}_1, \ldots, z^{\ast}_m) $ the set of weights is obtained by the following procedure: For each observation $(y_j,\bold{z}_j)$ in the learning data set $\{(y_i,\bold{z}_i)\}_{i=1\ldots,n}$ a weight is calculated by counting the number of trees in which $(y_j,\bold{z}_j)$ ends up in the same terminal node as the new set of covariates $\bold{z}^{\ast}$. 
$$
\omega^{forest}_i(\bold{z}^{\ast}) = \sum_{s=1}^T \sum_{b=1}^{B^s} I((\bold{z}_i \in \mathcal{B}^s_b) \land (\bold{z}^{\ast} \in \mathcal{B}^s_b))
$$
where $T$ is the number of trees of the forest, $B^s$ is the number of terminal nodes of the \textit{s}-th tree and $\mathcal{B}^s_b$ represents the \textit{b}-th terminal node of the \textit{s}-th tree.\\
This individual set of weights $\bold{\omega}^{forest}(\bold{z}^{\ast}) = (\omega^{forest}_1(\bold{z}^{\ast}), \ldots, \omega^{forest}_n(\bold{z}^{\ast}))$ for the new set of covariates $\bold{z}^{\ast}$ can now be included in the estimation process of the distribution parameters and in that way, leads to its individual parameter vector $\hat{\theta}(\bold{z}^{\ast})$.
$$
\hat{\theta}(\bold{z}^{\ast}) =  \max_{\theta \in \Theta} \sum_{i=1}^n \omega^{forest}_i(\bold{z}^{\ast}) \cdot \ell(\theta; y_i)
$$
While in a distributional tree the same distribution is specified for all sets of covariates being in the same node, a distributional forest predicts and fits a different distribution for each different set of covariates which is due to the individual weights calculated separately for each set of covariates.\\
\\
As this novel methodology combines the advantages of smooth and flexible parametric modeling with those of tree structured modeling it provides a model that can deal with a wide range of situations. Abrupt changes and non-additive effects can be captured by its tree structure while smooth effects can be modeled reasonably by combining various tree and smoothing the steps imposed by a single tree. As scores are considered for each distribution parameter (separate) effects in different parameters can be captured. At the same time, also interactions are detected automatically by splitting the data in a way that is sensitive to changes in either each of the parameters or combined in various parameters.\\
Even though many modeling situations demand for the use of a specific model providing the needed features and therefore being clearly advantageous and the best choice, it is often hard to tell in advance what kind of situation this is. In these cases a distributional forest is a recommendable compromise providing reasonable models and performing well without any prespecification or additional knowledge needed in advance.

@@

\subsection{Data}
\label{data}
The data set used for training and validation consists of observed daily precipitation sums provided by the National Hydrographical Service (\citealt{ehyd}) and numerical weather forecasts from the {U.S.} National Oceanic and Atmospheric Administration (NOAA).
Both, observations and weather forecasts are available from 1985 trough the year 2012, focusing on July only in this article.\\ 
Observations are available on a daily basis, each observation representing the total precipitation sum fallen within 24~hours observed at 6~UTC at 95 stations all over Tyrol and it's close surrounding. The observations are undergoing a rigorous quality check before made available to the public.  As covariates the second generation reforecast data set of the global ensemble forecast system (GEFS) is used (\citealt{hamill2013}).  This data set consists of an 11-member ensemble based on a fixed version of the numerical model and a horizontal grid-spacing of about $50~\times~50 \text{km}^2$ initialized at 0~UTC from December 1984 to present. Each of the 11 ensemble members uses slightly different perturbed initial conditions trying to predict the situation specific uncertainty of the atmospheric state.

@@

In particular, 14 basic forecast variables are considered with up to 12 variations of each of these variables being included as covariates. Since each forecast variable consists of an ensemble of individual predictions, the possible variations include the mean, the standard deviation, the minimum and the maximum over all ensemble members, noted as ensemble mean, ensemble std dev, ensemble minimum and ensemble maximum. All of these values can be calculated on the sums over 24 hours (6--6 UTC of the following day) or a 6 hour time window.\\
Another possible option is to consider the mean, minimum or maximum over 24 hours for each individual prediction and then take the ensemble mean or ensemble standard deviation, noted as ensemble mean of mean/minimal/maximal values or ensemble std dev of mean/minimal/maximal values respectively.\\
Table~\ref{covariates} shows all variables together with the number of variations and the provided variations listed in the last column. Altogether 82 covariates are available in this setting.\\
To remove large parts of the skewness of precipitation data, a power transformation has been used frequently (Box and Cox 1964). In literature, cubic (\citealt{stidd1973}) or square root (\citealt{hutchinson1998b}) transformations have often been suggested but may vary for different climatic zones or temporal aggregation periods. In \citealt{stauffer2017a} optimization methods have been applied to find the best transformation factor which resulted in a value of $\frac{1}{1.6}$. Since the data used for this optimization covers almost the same geographical region as considered in this article the same factor is chosen here for the power transformation of the response variable \texttt{tp} (total precipitation). The resulting transformed variable is expected to be normally distributed and left censored at censoring point 0. For simplicity the same transformation has been chosen for the variable \texttt{cape} (convective available potential energy).

%\newpage
\begin{table}[t!]
%\resizebox{1.4\textwidth}{!}{
\begin{minipage}{\textwidth}
\begin{tabular}{ l  c  l }
\hline
Basic covariates & {\#} & Variations\\
\hline
\textbf{tp}: total precipitation            & 12  & ensemble mean of sums over 24h, \\
\hspace*{\fill} power transformed (by $\frac{1}{1.6}$) &    & ensemble std dev of sums over 24h, \\ 
\textbf{cape}: convective                  &    & ensemble minimum of sums over 24h, \\
\hspace*{0.4cm} available potential energy &    & ensemble maximum of sums over 24h\\
\hspace*{\fill}  power transformed (by $\frac{1}{1.6}$)&    & \qquad all for 6--30 UTC\\
\hspace*{\fill}                            &    & ensemble mean of sums over 6h\\
\hspace*{\fill}                            &    & \qquad for 6--12, 12--18, 18--24, 24--30 UTC\\
\hspace*{\fill}                            &    & ensemble std dev of sums over 6h\\
\hspace*{\fill}                            &    & \qquad for 6--12, 12--18, 18--24, 24--30 UTC\\
\hline
\textbf{dswrf}: downwards short ware      & 6 & ensemble mean of mean values, \\
\hspace*{\fill} radiation flux (``sunshine'') &   & ensemble mean of minimal values,\\
\textbf{msl}: mean sea level pressure     &   &  ensemble mean of maximal values,\\
\textbf{pwat}: preciptable water          &   & ensemble std dev of mean values,\\
\textbf{tmax}: 2m maximum temperature     &   & ensemble std dev of minimal values,\\
\hspace*{1.15cm}                          &   & ensemble std dev of maximal values,\\
\textbf{tcolc}: total column-integrated   &   & \qquad all over 6--30 UTC \\
\hspace*{1.05cm} condensate               &   & \\
\textbf{t500}: temperature on 500 hPa     &   & \\
\hspace*{1cm}                             &   & \\
\textbf{t700}: temperature on 700 hPa     &   & \\
\hspace*{1cm}                             &   & \\
\textbf{t850}: temperature on 850 hPa     &   & \\
\hspace{1cm}                              &   & \\
\hline
\textbf{tdiff500850}: temperature         & 3 & ensemble mean of difference in mean,\\
\hspace*{\fill} difference 500 to 850 hPa &   & ensemble minimum of difference in mean,\\
\textbf{tdiff500700}: temperature         &   & ensemble maximum of difference in mean\\
\hspace*{\fill} difference 500 to 700 hPa &   & \qquad all over 6--30 UTC\\
\textbf{tdiff700850}: temperature         &   & \\
\hspace*{\fill} difference 700 to 850 hPa &   & \\
\hline
\textbf{msl{\_}diff}: mean sea level  & 1 & msl{\_}mean{\_}max - msl{\_}mean{\_}min\\
\hspace{1.65cm} difference            &   & \qquad over 6--30 UTC\\
\hline
\end{tabular}
\caption[Table caption text]{Basic covariates together with the number ({\#}) and the type of variations.}
\label{covariates}
\end{minipage} 
%}
\end{table}

\subsection{Models and Evaluation}
Next to distributional forests three other methods are applied to this data set and their results are compared based on CRPS values. One of the reference models is an EMOS models as this is a common choice in this type of setting and research field. The other two models are both GAMs, once in a prespecified form in terms of variable selection for each distribution parameters and once applying a boosting algorithm). 
The chosen methods and their specifications are listed in Table~\ref{model_specification}.
\begin{table}[t!]
%\resizebox{1.4\textwidth}{!}{
%\begin{minipage}{\textwidth}
\begin{center}
\begin{tabular}{ l  l   l   l }
\hline
Model & Type & Location ($\mu$) & Scale ($\log(\sigma)$) \\
\hline
\textbf{Forest} & recursive    & all & all \\ 
                & partitioning &     &     \\ 
\hline
\textbf{prespecified GAM} & spline  & tp{\_}mean, & tppow{\_}sprd,\\
                            & in each & tp{\_}max,  & dswrf{\_}sprd{\_}mean,\\
 & & tp{\_}mean1218 $\ast$ & tp{\_}sprd1218 $\ast$\\ 
 & & \quad cape{\_}mean1218, & \quad  cape{\_}mean1218,\\
 & & dswrf{\_}mean{\_}mean, & tcolc{\_}sprd{\_}mean,\\
 & & tcolc{\_}mean{\_}mean, & tdiff500850{\_}mean\\
 & & pwat{\_}mean{\_}mean, & \\
 & & tdiff500850{\_}mean, & \\
 & & msl{\_}diff & \\
\hline
\textbf{boosted GAM} & spline  & all & all \\
                       & in each &     &     \\
\hline
\textbf{EMOS} & linear & tp{\_}mean & tp{\_}spread \\
\hline
\end{tabular}
\caption[Table caption text]{Table of applied methods together with their type (way of including the covariates) and the included covariates for each distribution parameter.}
\label{model_specification}
%}
\end{center}
\end{table}

>>>>

\SweaveOpts{eval = TRUE}
\section{Introduction}
A typical application in meteorology is to improve probabilistic precipitation forecasts. The main interest lies in estimating the probability of precipitation as well as predicting the expected amount of precipitation. In this research area various methods have been developed and tested in the last decades. (Multiple) Linear models such as ensemble model output statistics (EMOS) were introduced more than a decade ago (\citealt{gneiting2005calibrated}) and are still common and widely used. Despite their simplicity and interpretability, the limitation to modeling only linear effects has risen the interest in applying more flexible models. This feature is provided by generalized additive models (GAM) which can also model non-linear effects. However, often only the expectation of the underlying distribution is estimated. To obtain probabilistic predictions, generalized additive models for location, scale, and shape (GAMLSS, \citealt{stasinopoulos2005}) offer a framework in which each distribution parameter can be modeled separately by its own GAM. In Figure~\ref{Devel_parmod} the development of parametric models starting from simple linear models (LM) and ranging to GAMLSS is illustrated.\\
These flexible parametric methods provide well fitting models when applied on precipitation data in a rather smooth and homogeneous setting, e.g. areas where no abrupt changes in meteorological/climatological events appear. However, for example in complex terrain in mountainous regions various factors can change rapidly and their effects are difficult to be represented by a smooth parametric model. In this situation a different approach which can deal with these abrupt changes would be to apply tree structured models such as regression trees (\citealt{loh2011classification}) or random forests (\citealt{breiman2001random}). Moreover, these models are particularly advantageous in case of many covariates with unknown effects or interactions as the variable selection is done automatically within the model fitting process. Therefore, no additional effort or knowledge for the specification of the model is required in advance. However, even though their main advantage of being able to model non-additive effects and interactions without any necessary prespecification can play to their strength, regression trees are not an adequate choice if (additional) smooth effects appear as well. In this situation a random forest provides a good compromise as it combines a set of trees such that the steps imposed by the trees are smoothed out in case of smooth effects while abrupt changes are still captured by the tree structure of the model.\\
For these reasons, regression trees and particularly random forests provide valuable features which can contribute to improving models for probabilistic forecasting. But, up to now, no general distributional framework is available for these methods. Therefore, we are going to introduce distributional trees and forests which embed the idea of flexible parametric modeling into the framework of regression trees and random forests. The developments in the field of tree-structured models together with the novel approaches of distributional trees and forests are illustrated in Figure~\ref{Devel_treeforest}.\\

@@

After explaining the novel methodology in Section~\ref{Methodology} of this paper the focus will then be put on an application in the field of probabilistic forecasting in Section~\ref{ProbabilisticForecasting}. The new method is applied on precipitation data in a mountainous region with a large number of numerical weather prediction quantities and the results are compared to those of other methods commonly used in this research area. In a first setting only one observation station is taken into consideration. The models are trained on data from 24 years and probabilistic predictions of the consecutive 4 years are validated. A brief outlook on the results of this setting is given by Figure~\ref{axams_onestation}. For the 24th of July of 2009, 2010, 2011 and 2012 the density functions of the distributions estimated by a forest model are plotted together with the corresponding observations. This illustrates that well fitting predicted expected values are provided by this model as they are close to the observed values. Beside the expected values the predictions provided by distributional models offer a wide range of information as they specify the whole distribution and therefore enable further methods of inference and analysis. In section \ref{ProbabilisticForecasting} this example will be further discussed and it will be shown that the predictions made by the forest model are reasonably well fitting, not only for the mean parameter of the distribution but also for its second parameter, the scale parameter, and therefore the whole distribution.

@@

where $\phi$ and $\Phi$ are the probability density function and the distribution function of the standard normal distribution $\mathcal{N}(0,1)$. Next to the censored normal distribution the censored logistic distribution is also often chosen for the modeling of precipitation amounts. In \citealt{Gebetsberger2017} this choice has been made in order to be able to deal with heavy tails which often appear in precipitation data. Another frequent application of probabilistic forecasting in meteorology is the modeling of near surface air temperature. In this case a normal distribution is often selected.\\

@@

In that way, a whole distribution is specified with all its features, including location, scale and shape. However, fitting one global model to the whole data set might be too much of a generalization and therefore not represent the data and its features in a reasonable way. Moreover, if covariates are available it is desirable to include them into the model as they might provide important additional information. In our application on precipitation data a wide range of covariates are available from a numerical ensemble prediction system considering precipitation, temperature, sea level pressure and other meteorological quantities which might be of interest and have an influence on the predictions. \\

@@

However, the restriction to additive models remains. Therefore, a different approach on how to include covariates that also allows for non-additive effects and interactions is considered in the next subsection of this article. In particular, the covariates can be used to separate the data into more homogeneous subgroups. Doing so after fitting a global model to the whole data set (without considering any covariates) and then fitting a local model to each subgroup should improve the model remarkably. This procedure of splitting a data set into subgroups can be done by applying a tree algorithm.\\

@@

While the clear structure of such a tree model is advantageous in interpreting and analyzing it, the abrupt changes are often too rough and impose steps on the model even if the true underlying effect is smooth. In these situations bagging can solve the problem by combining an ensemble of trees such that wrongly detected abrupt changes of the model are turned into almost smooth transitions. In that way, all kinds of effects can be modeled, which is is also one the of the main advantages of forest models (taking bagging one step further as the variance is reduced by sampling the covariates considered for the splitting process in each node). \\
\\
Common forest models such as random forests consist of trees only considering the mean of the distribution of the response variable. For that reason, changes in other parameters can only be detected if they are connected to changes in the location parameter, e.g. if they depend on the location parameter. Therefore, building a forest of distributional trees which can also capture separate effects in each of the distribution parameters leads to an improvement of the model. This valuable feature of detecting separate effects is also provided by GAMLSS as each parameter is modeled separately. However, in this case this separation can also be a disadvantage as it hinders the capturing of interactions unless they are specified explicitly by the user. These two cases provide once again a good reason to combine the two frameworks of flexible parametric modeling and tree structured models in order to profit from their advantages and compensate possible disadvantages. This is the pursued goal in introducing distributional forests.

\subsection{Distributional Forest}
\label{distforest}
A distributional forest consists of an ensemble of distributional trees $\{t_s\}_{s=1,\ldots,T}$. Each of these trees is built on a different data set which can either be a bootstrap or a subset of the original data set. Moreover, in each node a subset of the partitioning variables is chosen randomly as candidates for splitting variables. In that way the correlation among the resulting trees is reduced and therefore also the variance of the model (\citealt{breiman2001random}).\\
\\
Forest models differ mainly in the way how the trees are combined to make predictions. In this case the forest itself is only used to evaluate neighborhood relationships among the observations which can then be included in the fitting process.\\
Contrary to distributional trees, in distributional forests the whole learning data set can be used to fit the distributional model and specify the predicted distribution for a new set of covariates. This difference is represented in the weight vector as the tree weights $\bold{\omega}^{tree}$ containing only zeros and ones are now replaced by integer valued forest weights \\
$\bold{\omega}^{forest} = (\omega^{forest}_1, \ldots, \omega^{forest}_n)$.\\
For a new set of covariates $\bold{z}^{\ast} = (z^{\ast}_1, \ldots, z^{\ast}_m) $ the set of weights is obtained by the following procedure: For each tuple $(y_j,\bold{z}_j)$ in the learning data set $\{(y_i,\bold{z}_i)\}_{i=1\ldots,n}$ a weight is calculated by counting the number of trees in which $(y_j,\bold{z}_j)$ ends up in the same terminal node as the new set of covariates $\bold{z}^{\ast}$. 
$$
\omega^{forest}_i(\bold{z}^{\ast}) = \sum_{s=1}^T \sum_{b=1}^{B^s} I((\bold{z}_i \in \mathcal{B}^s_b) \land (\bold{z}^{\ast} \in \mathcal{B}^s_b))
$$
where $T$ is the number of trees of the forest, $B^s$ is the number of terminal nodes of the \textit{s}-th tree and $\mathcal{B}^s_b$ represents the \textit{b}-th terminal node of the \textit{s}-th tree.\\
This individual set of weights $\bold{\omega}^{forest}(\bold{z}^{\ast}) = (\omega^{forest}_1(\bold{z}^{\ast}), \ldots, \omega^{forest}_n(\bold{z}^{\ast}))$ for the new set of covariates $\bold{z}^{\ast}$ can now be included in the estimation process of the distribution parameters and in that way leads to its individual parameter vector $\hat{\theta}(\bold{z}^{\ast})$.
$$
\hat{\theta}(\bold{z}^{\ast}) =  \max_{\theta \in \Theta} \sum_{i=1}^n \omega^{forest}_i(\bold{z}^{\ast}) \cdot \ell(\theta; y_i)
$$
While in a distributional tree the same distribution is specified for all sets of covariates being in the same node, a distributional forest predicts and fits a different distribution for each different set of covariates due to the individual weights calculated separately for each set of covariates.\\
\\
As this novel methodology combines the advantages of smooth and flexible parametric modeling with those of tree structured modeling it provides a model that can deal with a wide range of situations. Abrupt changes and non-additive effects can be captured by its tree structure while smooth effects can be modeled reasonably well by combining various tree and smoothing the steps imposed by a single tree. As scores are considered for each distribution parameter (separate) effects in different parameters can be captured. At the same time, also interactions are detected automatically by splitting the data in a way that is sensitive to changes in either each of the parameters or combined in various parameters.\\
Even though many situations demand for the use of specificaly taylored models providing the needed features and therefore being clearly advantageous and the best choice, it is often hard to tell in advance which features will be required. In these cases a distributional forest is a recommendable compromise providing reasonable models and performing well without any prespecification or additional knowledge needed in advance.

@@

\subsection{Data}
\label{data}
The data set used for training and validation consists of observed daily precipitation sums provided by the National Hydrographical Service (\citealt{ehyd}) and numerical weather forecasts from the {U.S.} National Oceanic and Atmospheric Administration (NOAA).
Both, observations and weather forecasts are available from 1985 trough the year 2012, focusing on July only in this article.\\ 
Observations are available on a daily basis, each observation representing the total precipitation sum fallen within 24~hours observed at 6~UTC at 95 stations all over Tyrol and it's close surrounding. The observations are undergoing a rigorous quality check before made available to the public.  As covariates the second generation reforecast data set of the global ensemble forecast system (GEFS) is used (\citealt{hamill2013}).  This data set consists of an 11-member ensemble based on a fixed version of the numerical model and a horizontal grid-spacing of about $50~\times~50 \text{km}^2$ initialized daily at 0~UTC from December 1984 to present providing forecasts on 26 hourly temporal resolution. Each of the 11 ensemble members uses slightly different perturbed initial conditions trying to predict the situation specific uncertainty of the atmospheric state.

@@

In particular, 14 basic forecast variables are considered with up to 12 variations of each of these variables being included as covariates. Since each forecast variable consists of an ensemble of individual predictions, the possible variations include the mean, the standard deviation, the minimum and the maximum over all ensemble members, noted as ensemble mean, ensemble std dev, ensemble minimum and ensemble maximum. All of these values can be calculated on the sums over 24 hours (6--6 UTC of the following day) or a 6 hour time window.\\
Another possible option is to consider the mean, minimum or maximum over 24 hours for each individual prediction and then take the ensemble mean or ensemble standard deviation, noted as ensemble mean of mean/minimal/maximal values or ensemble std dev of mean/minimal/maximal values respectively.\\
Table~\ref{covariates} shows all variables together with the number of variations and the provided variations listed in the last column. Altogether 82 covariates are available in this setting.\\
To remove large parts of the skewness of precipitation data, a power transformation has been used frequently (Box and Cox 1964). In literature, cubic (\citealt{stidd1973}) or square root (\citealt{hutchinson1998b}) transformations have often been suggested but may vary for different climatic zones or temporal aggregation periods. In \citealt{stauffer2017a} optimization methods have been applied to find the best transformation factor which resulted in a value of $\frac{1}{1.6}$. Since the observation data set used for this optimization contains the observations considered in this article the same factor is chosen here for the power transformation of the response variable \texttt{tp} (total precipitation). The resulting transformed variable is expected to be normally distributed and left censored at censoring point 0. For simplicity the same transformation has been chosen for the variable \texttt{cape} (convective available potential energy).

%\newpage
\begin{table}[t!]
%\resizebox{1.4\textwidth}{!}{
\begin{minipage}{\textwidth}
\begin{tabular}{ l  c  l }
\hline
Basic covariates & {\#} & Variations\\
\hline
\textbf{tp}: total precipitation            & 12  & ensemble mean of sums over 24h, \\
\hspace*{\fill} power transformed (by $\frac{1}{1.6}$) &    & ensemble std dev of sums over 24h, \\ 
\textbf{cape}: convective                  &    & ensemble minimum of sums over 24h, \\
\hspace*{0.4cm} available potential energy &    & ensemble maximum of sums over 24h\\
\hspace*{\fill}  power transformed (by $\frac{1}{1.6}$)&    & \qquad all for 6--30 UTC\\
\hspace*{\fill}                            &    & ensemble mean of sums over 6h\\
\hspace*{\fill}                            &    & \qquad for 6--12, 12--18, 18--24, 24--30 UTC\\
\hspace*{\fill}                            &    & ensemble std dev of sums over 6h\\
\hspace*{\fill}                            &    & \qquad for 6--12, 12--18, 18--24, 24--30 UTC\\
\hline
\textbf{dswrf}: downwards short wave      & 6 & ensemble mean of mean values, \\
\hspace*{\fill} radiation flux (``sunshine'') &   & ensemble mean of minimal values,\\
\textbf{msl}: mean sea level pressure     &   &  ensemble mean of maximal values,\\
\textbf{pwat}: preciptable water          &   & ensemble std dev of mean values,\\
\textbf{tmax}: 2m maximum temperature     &   & ensemble std dev of minimal values,\\
\hspace*{1.15cm}                          &   & ensemble std dev of maximal values,\\
\textbf{tcolc}: total column-integrated   &   & \qquad all over 6--30 UTC \\
\hspace*{1.05cm} condensate               &   & \\
\textbf{t500}: temperature on 500 hPa     &   & \\
\hspace*{1cm}                             &   & \\
\textbf{t700}: temperature on 700 hPa     &   & \\
\hspace*{1cm}                             &   & \\
\textbf{t850}: temperature on 850 hPa     &   & \\
\hspace{1cm}                              &   & \\
\hline
\textbf{tdiff500850}: temperature         & 3 & ensemble mean of difference in mean,\\
\hspace*{\fill} difference 500 to 850 hPa &   & ensemble minimum of difference in mean,\\
\textbf{tdiff500700}: temperature         &   & ensemble maximum of difference in mean\\
\hspace*{\fill} difference 500 to 700 hPa &   & \qquad all over 6--30 UTC\\
\textbf{tdiff700850}: temperature         &   & \\
\hspace*{\fill} difference 700 to 850 hPa &   & \\
\hline
\textbf{msl{\_}diff}: mean sea level pressure & 1 & msl{\_}mean{\_}max - msl{\_}mean{\_}min\\
\hspace{1.45cm} difference            &   & \qquad over 6--30 UTC\\
\hline
\end{tabular}
\caption[Table caption text]{Basic covariates together with the number ({\#}) and the type of variations.}
\label{covariates}
\end{minipage} 
%}
\end{table}

\subsection{Models and Evaluation}
Next to distributional forests three other methods are applied to this data set and their results are compared based on the continuous ranked probability score (CRPS). One of the reference models is an EMOS model as this is a common choice in this type of setting and research field. The other two models are both GAMs, once in a prespecified form in terms of variable selection for each distribution parameter and once applying a boosting algorithm. 
The chosen methods and their specifications are listed in Table~\ref{model_specification}.
\begin{table}[t!]
%\resizebox{1.4\textwidth}{!}{
%\begin{minipage}{\textwidth}
\begin{center}
\begin{tabular}{ l  l   l   l }
\hline
Model & Type & Location ($\mu$) & Scale ($\log(\sigma)$) \\
\hline
\textbf{Forest} & recursive    & all & all \\ 
                & partitioning &     &     \\ 
\hline
\textbf{prespecified GAM} & spline  & tp{\_}mean, & tppow{\_}sprd,\\
                            & in each & tp{\_}max,  & dswrf{\_}sprd{\_}mean,\\
 & & tp{\_}mean1218 $\ast$ & tp{\_}sprd1218 $\ast$\\ 
 & & \quad cape{\_}mean1218, & \quad  cape{\_}mean1218,\\
 & & dswrf{\_}mean{\_}mean, & tcolc{\_}sprd{\_}mean,\\
 & & tcolc{\_}mean{\_}mean, & tdiff500850{\_}mean\\
 & & pwat{\_}mean{\_}mean, & \\
 & & tdiff500850{\_}mean, & \\
 & & msl{\_}diff & \\
\hline
\textbf{boosted GAM} & spline  & all & all \\
                       & in each &     &     \\
\hline
\textbf{EMOS} & linear & tp{\_}mean & tp{\_}spread \\
\hline
\end{tabular}
\caption[Table caption text]{Table of applied methods together with their type (way of including the covariates) and the included covariates for each distribution parameter.}
\label{model_specification}
%}
\end{center}
\end{table}

For the prespecified GAM model the covariates were selected manually based on meteorological knowledge and experience. Additionally, interactions are specified in advance, once between \texttt{tp{\_}mean1218} and \texttt{cape{\_}mean1218} for the location parameter and once between \texttt{tp{\_}sprd1218} and \texttt{cape{\_}mean1218} for the scale parameter. These interactions might capture events with a high potential for moderate or severe thunderstorms with large amounts of precipitation. By putting extra effort in the specification of the formulas the prespecified GAM model is provided with additional information and is therefore expected to profit from this advantage.\\
In the EMOS model the logarithmic function is chosen as the link function for the scale parameter. This choice has been made based on the results of \citealt{Gebetsberger2017} where different link functions (identity, quadratic and logarithmic function) have been compared. Moreover, testing the identity, quadratic and logarithmic function as link functions in this setting supported this choice as the logarithmic function lead to the best performance of the EMOS model in this application.\\
\\
In order to compare the performance of the chosen methods, the continuous ranked probability score (CRPS) is calculated for each model. The CRPS measures the accuracy of a probabilistic model. For one observation $y$ with the predicted distribution function $F$ the CRPS value is calculated by
$$
\text{CRPS}(y, F) = \int_{-\infty}^\infty (F(z) - \mathbf{1}(y \leq z))^2 dz
$$
where \(\mathbf{1}\) is the indicator function. \\
In this application the CRPS is evaluated out of sample on a data set that has not been used to build the model. The mean CRPS over thi data set is used for validation. Since the CRPS can be seen as a generalization of the mean absolute error, a low value indicates a good fit. In applications, one big advantage in terms of interpretability is that the resulting value is on the scale of the observation. \\
For the purpose of comparing the performance of a distributional forest to those of other models the so-called skills score can be calculated with respect to the CRPS and one model being chosen as a reference. In that way we obtain a measure of how much a method improves the model performance in terms of CRPS compared to the reference method.
$$
\text{SkillsScore}_{\text{method}} = 1 - \frac{\text{CRPS}_{\text{method}}}{\text{CRPS}_{\text{reference}}}
$$
